\ProjectName{}: Data Imputation as a Query Optimization

### 2017-02-23 paper/experiments.tex
paper: Update queries and descriptions in experiments sec

\subsection{Queries}
We collected a set of queries (Table~\ref{fig:queries}) that we think are both representative, in that they could reasonably be written by a user in the course of data analysis, and interesting to plan.

The queries on the CDC NHANES data consist not only of projections and selections,
but also interesting joins and aggregations. Our aim was to craft meaningful queries
that would provide performance figures relevant to practitioners who may use
similar datasets.

Queries 1-3 are on the CDC data. Query~\ref{q1} calculates the average weight for users based on their income data.
To do so, it joins demographics and physical examination data. Query~\ref{q2}
compares cholesterol levels for individuals who are at both ends of the income
spectrum and who satisfy a minimum weight requirement. Query~\ref{q3}
extracts the maximum blood lead levels for subjects who satisfy a particular
waist circumference requirement.

Queries 4-7 are on the FCC data. In Query~\ref{q4}, we calculate the average income for survey participants,
based on their bootcamp attendance. Query~\ref{q5} estimates the average
age of women from the United States who participated. Query~\ref{q6} 
calculates the average amount of money survey participants with student debt destined to
learning based on their school degree. Finally, in Query~\ref{q7}, we join the FCC data with a reference table provided
by the World Bank which summarizes GDP per Capita across various
countries\cite{worldbank-data}. The query calculates the average GDP-per-capita for participants based on whether or not they
attended a bootcamp.

@@

%\todobox{actually run experiments and rewrite this section}{
All experiments were run on a Macbook Pro with two 2.9 GHz Intel core i5 processors, 16 GB of main memory,  running OS X El Capitan.

@@

Figure~\ref{fig:plantimes} provides a summary of the planning times for each of the queries.  We exclude the planning time for queries that impute at base table, as that
requires no planning. On average, planning constituted 38 percent of total runtime for plans generated by \ProjectName{}.

@@

Table~\ref{table:smape} shows the Symmetric-Mean-Absolute-Percentage-Error (SMAPE) for \ProjectName{}'s query results when compared to 
running imputation on the base tables and executing the query on the
cleaned data. Each query was run 10 iterations in both settings and then results from the two approaches were paired up and compared tuple-wise. We average tuple-wise
absolute percentage deviations within each iteration of a query, and we report this value averaged over all iterations.
We can see that optimizing for quality indeed reduces the SMAPE of query results. In general, the SMAPE relative to the
base-imputation approach are low in all cases, indicating that on-the-fly imputation produces similar results to
imputation at the base tables.

@@

In many cases, applying the imputation step at the base table simply is not an option. To illustrate the increasing difficulty of such an approach
as datasets scale, we used a simple aggregate query over the ACS dataset in our final experiment(Figure~\ref{query-acs}). Applying the imputation operation to the base table
is extremely expensive, in our case the process did not terminate within 20 minutes. In contrast, \ProjectName{} executes a quality-optimized version
in 9 seconds and a runtime-optimized version in 1 second. This highlights the potential benefit of using our system for early data exploration.

\begin{figure}
\begin{lstlisting}
SELECT AVG(c0) FROM acs_dirty;
\end{lstlisting}
\caption{A simple aggregate query on ACS data. Imputing the base table is undesirable (did not terminate within 20min). \ProjectName's
quality-optimized and performance-optimized query execute in 9 seconds and 1 second, respectively.
}
\label{query-acs}
\end{figure}

>>>>

\subsection{Queries}
We collect a set of queries (Table~\ref{fig:queries}) that we think are both representative, in that they could reasonably be written by a user in the course of data analysis, and interesting to plan.

The queries on the CDC NHANES data consist not only of projections and selections,
but also interesting joins and aggregates. Our aim was to craft meaningful queries
that would provide performance figures relevant to practitioners who may use
similar datasets.

Queries 1-5 are on the CDC data. In Query~\ref{q1}, we calculate the average height for users based on their income data. In Query~\ref{q2},
we compare cholesterol levels for individuals at either tail of the income
spectrum and above a certain weight. In Query~\ref{q3},
we extract the maximum blood lead levels for children under the age of 6 years. In Query ~\ref{q4}, we calculate the average systolic blood pressure, by gender, for subjects with a body mass index indicating obesity. In Query ~\ref{q5}, we extract the age (in months), creatine level, and head circumference, for subjects with high creatine levels.

% Perhaps more school makes your head bigger :)

Queries 6-10 are on the FCC data. In Query~\ref{q6}, we calculate the average income for survey participants,
based on their bootcamp attendance. In Query~\ref{q7} estimates the average
age of women from the United States who participated. In Query~\ref{q8}, we 
calculate the average amount of money survey participants with student debt destined to
learning based on their school degree. In Query~\ref{q9}, we join the FCC data with a reference table provided
by the World Bank which summarizes GDP per Capita across various
countries\cite{worldbank-data}. The query calculates the average GDP-per-capita for participants based on whether or not they
attended a bootcamp. Finally, in Query ~\ref{q10}, we identify individuals from countries with GDP per-capita of \$5000 or
less who earn \$100,000 or more post-bootcamp.

@@

%\todobox{actually run experiments and rewrite this section}{
All experiments were run on a single AWS EC2 c4.xlarge instance, with 4 2.9 GHz Intel Xeon E5-2666 v3 virtual CPUs and 7.5 GiB of main memory on Debian Linux.

@@

%TODO MJS
Figure~\ref{fig:plantimes} provides a summary of the planning times for each of the queries.  We exclude the planning time for queries that impute at base table, as that
requires no planning. On average, planning constituted 38 percent of total runtime for plans generated by \ProjectName{}.

@@

Table~\ref{table:smape} shows the Symmetric-Mean-Absolute-Percentage-Error (SMAPE) for \ProjectName{}'s query results when compared to 
running imputation on the base tables and executing the query on the
cleaned data. Each query was run 10 %TODO MJS
iterations in both settings and then results from the two approaches were paired up and compared tuple-wise. We average tuple-wise
absolute percentage deviations within each iteration of a query, and we report this value averaged over all iterations.
We can see that optimizing for quality indeed reduces the SMAPE of query results. In general, the SMAPE relative to the
base-imputation approach are low in all cases, indicating that on-the-fly imputation produces similar results to
imputation at the base tables.

@@

% TODO MJS
In many cases, applying the imputation step at the base table simply is not an option. To illustrate the increasing difficulty of such an approach
as datasets scale, we used a simple aggregate query over the ACS dataset in our final experiment(Figure~\ref{query-acs}). Applying the imputation operation to the base table
is extremely expensive, in our case the process did not terminate within 20 minutes. In contrast, \ProjectName{} executes a quality-optimized version
in 9 seconds and a runtime-optimized version in 1 second. This highlights the potential benefit of using our system for early data exploration.

\begin{figure}
\begin{lstlisting}
SELECT AVG(c0) FROM acs_dirty;
\end{lstlisting}
\caption{A simple aggregate query on ACS data. Imputing the base table is undesirable (did not terminate within 20min). \ProjectName's
quality-optimized and performance-optimized query execute in 9 seconds and 1 second, respectively.
}%TODO MJS
\label{query-acs}
\end{figure}
